---
title: "REDD_LinearModels"
output: pdf_document
date: "2025-08-15"
---

Title: REDD+ Linear Models After Matching (Scorecard LMs)
Author: Kelly Ong

Description:
This script runs project-level fixed-effects linear models on matched pixel 
datasets for five score-card variables:
1) forestInt   
2) forestFrag   
3) carbon_flux
4) canopy_height  
5) biodIntact2015

Workflow:
(A) Read REDD+ project list.
(B) Exclude UID–biome combinations with any post-match SMD > 0.25.
(C) For each comparison, load matched pixels (treated vs control),
  require >= 30 complete pairs, build the 4-level "group", and
  fit: outcome ~ group + [unbalanced covariates] + factor(pair_id).
(D) Save per-UID coefficient tables and a combined summary CSV.
(E) Run sensitivity analysis for unobserved confounding (sensemakr)

Design choices:
- Pair fixed effects via factor(pair_id) respect 1:1 matching.
- Only covariates that remain unbalanced within pairs (paired t-test p<0.05)
enter the LM as confounders. This guards against overfitting and collinearity.
- Projects with poor balance (any SMD > 0.25) are excluded a priori.

Outputs (per comparison × score-card variable):
- output/LM_*/<var>/lm_summary_<UID>.csv (per-project coefficients)
- output/LM_*/<var>/lm_summary_combined_<var>.csv
- output/LM_*/<var>/LM_skipped_REDD_sites.csv (reasons for skips)

Notes:
Coefficients on the 'group<treated>' term are interpreted as the
average within-pair difference (treated vs control) in the outcome,
conditional on pair fixed effects and any unbalanced covariates.

# 0. Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)

# Set the working directory where all datasets are located
wd <- "~/Documents/Ecological Integrity of Avoided Deforestation Projects/data/"

# Load required packages
library(readr)
library(sf)
library(dplyr)
library(tidyr)
library(readr)
library(rlang)
library(purrr)
library(sensemakr)

# Small helper
msg <- function(...) cat("[", format(Sys.time()), "] ", ..., "\n", sep = "")
```

# 1. Read core data

```{r}
# REDD+ project polygons (with REDD UID & biome info)
redd <- st_read(paste0(wd, "REDD.shp")) # REDD project boundaries

# Project-specific pixel resolution (km)
area <- read.csv(file.path(wd, "area.csv"))

# Candidate project IDs
file_ids <- area$UID
```

# 2. Exclude UID–biome combinations with any SMD > 0.25

```{r}
# Rationale: enforce covariate balance before modeling outcomes.
smd_dir <- paste0(wd, "output/NN_output_SMD_results_mahvars/")

redd_high_smd <- tibble::tibble(UID = character(), BIOME_NUM = numeric())
smd_file_issues <- tibble::tibble(UID = character(), Reason = character())

for (i in seq_len(nrow(redd))) {
  uid  <- redd$UID[i]
  bio  <- redd$BIOME_NUM[i]
  fpth <- paste0(smd_dir, sprintf("NN_SMD_%s_biome_%s.csv", uid, bio))

  if (!file.exists(fpth)) {
    smd_file_issues <- dplyr::bind_rows(smd_file_issues,
                                        tibble::tibble(UID = uid, Reason = "SMD file not found"))
    next
  }

  smd_row <- read.csv(fpth)
  if ("Covariate" %in% names(smd_row))
    smd_row <- smd_row[!(smd_row$Covariate == "distance"), , drop = FALSE]

  # Check if there are any SMDs>0.25
  if (any(abs(smd_row$SMD) > 0.25, na.rm = TRUE)) {
    redd_high_smd <- dplyr::bind_rows(redd_high_smd,
                                      tibble::tibble(UID = uid, BIOME_NUM = as.numeric(bio)))
  }
}

redd_clean <- redd %>%
  dplyr::anti_join(redd_high_smd, by = c("UID", "BIOME_NUM"))

msg("After SMD>0.25 filter:", nrow(redd_clean), "rows (UID–biome).")
file_ids_clean <- unique(redd_clean$UID)
msg("Distinct UIDs retained:", length(file_ids_clean))

# (Optional) record SMD file issues for transparency
if (!dir.exists(file.path(wd, "output"))) dir.create(file.path(wd, "output"))
readr::write_csv(smd_file_issues, file.path(wd, "output", "LM_prechecks_SMD_missing.csv"))
```

# 3. Helpers for LM stage

```{r}
# Matching covariates to test for residual imbalance within pairs.
matching_vars <- c(
  "aboveground_biomass_2000",
  "precipitation",
  "elevation",
  "slope",
  "human_footprint",
  "gdp",
  "population_density",
  "accessibility"
)

# (i) Paired t-test within matched pairs (REDD vs non-REDD) for a covariate.
#     Returns p-value; only covariates with p<0.05 are added to LM.
paired_test <- function(df, var){
  if (!all(c("pair_id","inside_REDD") %in% names(df))) return(NA_real_)
  tmp <- df[!is.na(df[[var]]), c(var, "inside_REDD", "pair_id")]
  # Retain pairs with one treated (1) and one control (0)
  pair_tab <- table(tmp$pair_id, tmp$inside_REDD)
  good <- rownames(pair_tab)[rowSums(pair_tab == 1) == 2]
  if (!length(good)) return(NA_real_)
  tmp <- tmp[tmp$pair_id %in% good, , drop = FALSE]
  if (length(unique(tmp$pair_id)) < 2) return(NA_real_)

  wide <- tidyr::pivot_wider(
    tmp, id_cols = pair_id,
    names_from = inside_REDD, values_from = !!sym(var),
    names_prefix = "treat_"
  )
  diffs <- wide$treat_1 - wide$treat_0
  if (length(diffs) < 2 || all(is.na(diffs))) return(NA_real_)
  stats::t.test(diffs, mu = 0)$p.value
}

# (ii) Loader for matched data for a UID.
#      Prefers a combined file (if you created one), else binds all per-biome files
#      that passed the SMD filter.
load_matched_for_uid <- function(uid, redd_clean, wd){
  dir_raw  <- file.path(wd, "output/NN_output_matched_data_mahvars")
  dir_comb <- file.path(wd, "output/NN_output_matched_data_mahvars_combined")

  comb_path <- file.path(dir_comb, sprintf("NN_matched_data_%s_combined.csv", uid))
  if (file.exists(comb_path)) {
    msg("   » Using combined file for ", uid)
    return(readr::read_csv(comb_path, show_col_types = FALSE))
  }

  biomes <- redd_clean$BIOME_NUM[redd_clean$UID == uid]
  candidates <- unique(file.path(dir_raw, sprintf("NN_matched_data_%s_biome_%s.csv", uid, biomes)))
  exists <- candidates[file.exists(candidates)]

  if (!length(exists)) {
    exists <- list.files(dir_raw,
                         pattern = sprintf("^NN_matched_data_%s_biome_\\d+\\.csv$", uid),
                         full.names = TRUE)
  }
  if (!length(exists)) return(NULL)

  msg("   » Binding ", length(exists), " biome file(s) for ", uid)
  suppressMessages(purrr::map_dfr(exists, ~ readr::read_csv(.x, show_col_types = FALSE)))
}

# (iii) Extract the group (treatment) effect from a fitted model.
#       Robust to additional covariates: matches the coefficient by its name.
extract_group_effect <- function(id, model, data, control_label){
  if (is.null(model) || is.null(data)) {
    return(tibble::tibble(UID=id, term=NA_character_, Estimate=NA_real_, StdError=NA_real_,
                          tvalue=NA_real_, pvalue=NA_real_))
  }
  # Ensure two levels only (control + treatment)
  glevels <- levels(droplevels(data$group))
  treated_level <- setdiff(glevels, control_label)
  # Fallbacks if levels are messy
  trm <- if (length(treated_level) == 1) paste0("group", treated_level) else NA_character_

  cf <- summary(model)$coefficients %>% as.data.frame()
  names(cf) <- sub("Std\\. Error", "StdError", names(cf))
  names(cf) <- sub("t value", "tvalue", names(cf))
  names(cf) <- sub("Pr\\(>\\|t\\|\\)", "pvalue", names(cf))

  if (!is.na(trm) && (trm %in% rownames(cf))) {
    return(tibble::tibble(
      UID      = id,
      term     = trm,
      Estimate = cf[trm, "Estimate"],
      StdError = cf[trm, "StdError"],
      tvalue   = cf[trm, "tvalue"],
      pvalue   = cf[trm, "pvalue"]
    ))
  } else {
    # Last resort: pick the single row that starts with "group"
    rn <- rownames(cf)
    grp_rows <- rn[grepl("^group", rn)]
    if (length(grp_rows) == 1) {
      trm <- grp_rows[1]
      return(tibble::tibble(
        UID      = id,
        term     = trm,
        Estimate = cf[trm, "Estimate"],
        StdError = cf[trm, "StdError"],
        tvalue   = cf[trm, "tvalue"],
        pvalue   = cf[trm, "pvalue"]
      ))
    }
  }
  tibble::tibble(UID = id, 
                 term = NA_character_, 
                 Estimate = NA_real_, 
                 StdError = NA_real_, 
                 tvalue = NA_real_, 
                 pvalue = NA_real_)
}

```

# 4. LM runner (one variable × one comparison)

```{r}
# Formula per-UID:  outcome ~ group + (unbalanced covariates) + factor(pair_id)
# Requirements:
# 1. >= 30 complete pairs with non-NA outcome
# 2. matched_data contains group flags and pair_id
run_scorecard_LM <- function(file_ids,
                             area,          # kept for interface consistency
                             control,       # e.g., "No PA, No REDD"
                             treatment,     # e.g., "No PA, REDD"
                             scorecard_var, # one of the five variables
                             output_dir,    # directory to write outputs
                             wd) {

  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

  lm_model_list      <- list()
  filtered_data_list <- list()
  skipped_ids <- tibble::tibble(UID = character(), Reason = character())

  for (id in file_ids) {
    msg("──────────────────────────────────────────")
    msg("Processing REDD site: ", id)

    matched_data <- load_matched_for_uid(id, redd_clean, wd)
    if (is.null(matched_data)) {
      skipped_ids <- dplyr::bind_rows(skipped_ids,
                                      tibble::tibble(UID = id, Reason = "Matched data file not found"))
      next
    }

    # Outcome availability checks
    if (!(scorecard_var %in% names(matched_data))) {
      skipped_ids <- dplyr::bind_rows(skipped_ids,
                                      tibble::tibble(UID = id, Reason = paste("Missing variable:", scorecard_var)))
      next
    }
    if (all(is.na(matched_data[[scorecard_var]]))) {
      skipped_ids <- dplyr::bind_rows(skipped_ids,
                                      tibble::tibble(UID = id, Reason = paste("All NA in", scorecard_var)))
      next
    }
    if (all(matched_data[[scorecard_var]] == 0, na.rm = TRUE)) {
      skipped_ids <- dplyr::bind_rows(skipped_ids,
                                      tibble::tibble(UID = id, Reason = paste("All zero in", scorecard_var)))
      next
    }

    # Build 4-category group; filter to the two levels of interest
    matched_data$group <- with(matched_data, ifelse(
      inside_REDD==1 & inside_protected_area==0, "No PA, REDD",
      ifelse(inside_REDD==0 & inside_protected_area==0, "No PA, No REDD",
        ifelse(inside_REDD==0 & inside_protected_area==1, "PA, No REDD",
          ifelse(inside_REDD==1 & inside_protected_area==1, "PA, REDD", NA)))))
    matched_data$group <- factor(matched_data$group)

    group_labels <- c(treatment, control)
    matched_data <- matched_data[matched_data$group %in% group_labels, , drop = FALSE]

    # Require >= 30 complete pairs with non-NA outcome
    if (!all(c("pair_id","group") %in% names(matched_data))) {
      skipped_ids <- bind_rows(skipped_ids, tibble::tibble(UID = id, Reason = "Missing pair_id or group"))
      next
    }

    valid_pairs <- matched_data %>%
      dplyr::filter(!is.na(.data[[scorecard_var]])) %>%
      dplyr::group_by(pair_id) %>%
      dplyr::filter(dplyr::n_distinct(group) == 2) %>%
      dplyr::ungroup() %>%
      dplyr::pull(pair_id) %>%
      unique()

    if (length(valid_pairs) < 30) {
      skipped_ids <- dplyr::bind_rows(skipped_ids,
                                      tibble::tibble(UID = id, Reason = paste0("<30 matched pairs with non-NA ", scorecard_var)))
      next
    }

    filtered_data <- matched_data %>%
      dplyr::filter(pair_id %in% valid_pairs) %>%
      dplyr::mutate(
        group = droplevels(group),                # drop unused levels → 2 levels only
        inside_REDD = as.factor(inside_REDD),
        REDD = as.factor(REDD),
        inside_protected_area = as.factor(inside_protected_area)
      )
    filtered_data$group <- stats::relevel(filtered_data$group, ref = control)

    # Select residual confounders (unbalanced within pairs)
    unbalanced_vars <- character(0)
    for (v in matching_vars) {
      if (v %in% names(filtered_data)) {
        pval <- paired_test(filtered_data, v)
        if (!is.na(pval) && pval < 0.05) unbalanced_vars <- c(unbalanced_vars, v)
      }
    }
    if (length(unbalanced_vars)) {
      msg("   » Confounders added to LM: ", paste(unbalanced_vars, collapse = ", "))
    } else {
      msg("   » No matching variable p<0.05; LM uses none.")
    }

    # Fit LM with pair fixed effects
    rhs <- c("group", unbalanced_vars, "factor(pair_id)")
    fmla <- stats::as.formula(paste(scorecard_var, "~", paste(rhs, collapse = " + ")))
    model <- tryCatch(stats::lm(fmla, data = filtered_data), error = function(e) {
      msg("   » LM error: ", e$message); return(NULL)
    })
    if (is.null(model)) {
      skipped_ids <- bind_rows(skipped_ids, tibble::tibble(UID = id, Reason = "LM error"))
      next
    }

    # Save per-UID coefficient table
    cf <- summary(model)$coefficients %>% as.data.frame()
    names(cf) <- sub("Std\\. Error", "StdError", names(cf))
    names(cf) <- sub("t value", "tvalue", names(cf))
    names(cf) <- sub("Pr\\(>\\|t\\|\\)", "pvalue", names(cf))
    cf$term <- rownames(cf); rownames(cf) <- NULL

    readr::write_csv(
      cf[, c("term","Estimate","StdError","tvalue","pvalue")],
      file.path(output_dir, sprintf("lm_summary_%s.csv", id))
    )

    lm_model_list[[id]]      <- model
    filtered_data_list[[id]] <- filtered_data
    msg("   » Saved outputs for ", id)
  }

  # Combined summary: the single treatment-vs-control coefficient per UID
  lm_summary_rows <- purrr::map_dfr(
    file_ids,
    ~ extract_group_effect(.x, lm_model_list[[.x]], filtered_data_list[[.x]], control_label = control)
  )

  readr::write_csv(
    lm_summary_rows,
    file.path(output_dir, sprintf("lm_summary_combined_%s.csv", scorecard_var))
  )

  readr::write_csv(
    skipped_ids,
    file.path(output_dir, "LM_skipped_REDD_sites.csv")
  )

  invisible(list(models  = lm_model_list,
                 matched = filtered_data_list,
                 skipped = skipped_ids,
                 combined= lm_summary_rows))
}
```

# 5. Run the four comparisons (for all five variables)

```{r}
# Comparisons:
# (1) 'REDD, No PA' vs 'No REDD, No PA'
# (2) 'REDD, No PA' vs 'No REDD, PA'
# (3) 'REDD, PA'    vs 'No REDD, No PA'
# (4) 'REDD, PA'    vs 'No REDD, PA'

outdir <- function(...) file.path(wd, "output", ...)

# 5.1 REDD, No PA vs No REDD, No PA
forestInt_1 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "forestInt",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_noPA","forestInt"),
  wd = wd)

forestFrag_1 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "forestFrag",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_noPA","forestFrag"),
  wd = wd) 

biod_intact_1 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "biodIntact2015",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_noPA","biodIntact2015"),
  wd = wd)

canopyheight_1 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "canopy_height",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_noPA","canopy_height"),
  wd = wd)

carbon_flux_1 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "carbon_flux",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_noPA","carbon_flux"),
  wd = wd)

# 5.2 REDD, No PA vs No REDD, PA
forestInt_2 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "forestInt",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_PA","forestInt"),
  wd = wd)

forestFrag_2 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "forestFrag",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_PA","forestFrag"),
  wd = wd)

biod_intact_2 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "biodIntact2015",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_PA","biodIntact2015"),
  wd = wd)

canopyheight_2 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "canopy_height",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_PA","canopy_height"),
  wd = wd)

carbon_flux_2 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "No PA, REDD",
  scorecard_var = "carbon_flux",
  output_dir = outdir("LM_summary/REDD_noPA_vs_noREDD_PA","carbon_flux"),
  wd = wd)

# 5.3 REDD, PA vs No REDD, No PA
forestInt_3 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "forestInt",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_noPA","forestInt"),
  wd = wd)

forestFrag_3 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "forestFrag",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_noPA","forestFrag"),
  wd = wd)

biod_intact_3 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "biodIntact2015",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_noPA","biodIntact2015"),
  wd = wd)

canopyheight_3 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "canopy_height",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_noPA","canopy_height"),
  wd = wd)

carbon_flux_3 <- run_scorecard_LM(file_ids_clean, area,
  control = "No PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "carbon_flux",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_noPA","carbon_flux"),
  wd = wd)

# 5.4 REDD, PA vs No REDD, PA 
forestInt_4 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "forestInt",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_PA","forestInt"),
  wd = wd)

forestFrag_4 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "forestFrag",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_PA","forestFrag"),
  wd = wd)

biod_intact_4 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "biodIntact2015",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_PA","biodIntact2015"),
  wd = wd)

canopyheight_4 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "canopy_height",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_PA","canopy_height"),
  wd = wd)

carbon_flux_4 <- run_scorecard_LM(file_ids_clean, area,
  control = "PA, No REDD", treatment = "PA, REDD",
  scorecard_var = "carbon_flux",
  output_dir = outdir("LM_summary/REDD_PA_vs_noREDD_PA","carbon_flux"),
  wd = wd)
```

# 6. Standardizing effect sizes for plotting

```{r}
# Helpers
safe_div <- function(num, den) {
  ifelse(is.finite(den) & den > 0, num / den, NA_real_)
}

get_max_valid <- function(x, use_abs = FALSE) {
  if (use_abs) x <- abs(x)
  x_no_na <- x[!is.na(x)]
  if (length(x_no_na) == 0) return(NA_real_)  # all NA
  if (all(x_no_na == 0)) return(NA_real_)  # all zeros (or 0+NA)
  max(x, na.rm = TRUE)
}

# 6.1 Read the linear model summaries for each comparison
read_panel <- function(panel_dir) {
  
  base <- paste0(
    wd,
    "output/",
    "LM_summary/",
    panel_dir
  )

  # Read each metric's combined summary
  lst <- list(
    forestInt      = read.csv(paste0(base, "forestInt/lm_summary_combined_forestInt.csv")),
    forestFrag     = read.csv(paste0(base, "forestFrag/lm_summary_combined_forestFrag.csv")),
    biodIntact2015 = read.csv(paste0(base, "biodIntact2015/lm_summary_combined_biodIntact2015.csv")),
    canopy_height  = read.csv(paste0(base, "canopy_height/lm_summary_combined_canopy_height.csv")),
    carbon_flux    = read.csv(paste0(base, "carbon_flux/lm_summary_combined_carbon_flux.csv"))
  )

  # Bind and standardize sign (flip signs for fragmentation & carbon flux)
  df <- dplyr::bind_rows(
    transform(lst$forestInt, variable = "forestInt"),
    transform(lst$forestFrag, variable = "forestFrag"),
    transform(lst$biodIntact2015, variable = "biodIntact2015"),
    transform(lst$canopy_height, variable = "canopy_height"),
    transform(lst$carbon_flux, variable = "carbon_flux")
  ) %>%
    dplyr::mutate(
      Estimate = ifelse(variable %in% c("forestFrag", "carbon_flux"), -Estimate, Estimate)
    )

  return(df)
}

panel_map <- data.frame(
  panel_dir = c(
    "REDD_noPA_vs_noREDD_noPA/",
    "REDD_noPA_vs_noREDD_PA/",
    "REDD_PA_vs_noREDD_noPA/",
    "REDD_PA_vs_noREDD_PA/"),
  panel_lab = c(
    "REDD (No PA) vs No REDD (No PA)",
    "REDD (No PA) vs No REDD (PA)",
    "REDD (PA) vs No REDD (No PA)",
    "REDD (PA) vs No REDD (PA)"
  )
)

# Build combined dataset
effects_df <- dplyr::bind_rows(
  transform(read_panel(panel_map$panel_dir[1]), panel = panel_map$panel_lab[1]),
  transform(read_panel(panel_map$panel_dir[2]), panel = panel_map$panel_lab[2]),
  transform(read_panel(panel_map$panel_dir[3]), panel = panel_map$panel_lab[3]),
  transform(read_panel(panel_map$panel_dir[4]), panel = panel_map$panel_lab[4])
)

# 6.2 Compute per-UID maxima from matched pixels
std_skipped <- data.frame(UID = character(), Reason = character())

max_list <- vector("list", length(file_ids_clean))
names(max_list) <- file_ids_clean

for (i in seq_along(file_ids_clean)) {
  uid <- file_ids_clean[i]
  md  <- load_matched_for_uid(uid, redd_clean, wd) # Reuse the loader from above

  if (is.null(md)) {
    std_skipped <- bind_rows(std_skipped, data.frame(UID = uid, Reason = "Matched data not found for maxima"))
    max_list[[i]] <- data.frame(
      UID = uid,
      forestInt = NA_real_, forestFrag = NA_real_, biodIntact2015 = NA_real_,
      canopy_height = NA_real_, carbon_flux = NA_real_
    )
    next
  }

  max_list[[i]] <- data.frame(
    UID            = uid,
    forestInt      = get_max_valid(md$forestInt),
    forestFrag     = get_max_valid(md$forestFrag),
    biodIntact2015 = get_max_valid(md$biodIntact2015),
    canopy_height  = get_max_valid(md$canopy_height),
    carbon_flux    = get_max_valid(md$carbon_flux, use_abs = TRUE) # absolute max for flux
  )
}

max_df <- bind_rows(max_list) %>%
  pivot_longer(-UID, names_to = "variable", values_to = "max_indicator_value")

# 6.3 Standardize: std_treatment_effect = Estimate / max
std_df <- effects_df %>%
  left_join(max_df, by = c("UID", "variable")) %>%
  mutate(
    std_treatment_effect = safe_div(Estimate, max_indicator_value),
    std_treatment_effect_zeroed = ifelse(pvalue >= 0.05, 0, std_treatment_effect)  # (optional) zero-out non-sig
  )

# Save consolidated and per-panel outputs
dir.create(file.path(wd, "output", "LM_standardized"), showWarnings = FALSE)

write.csv(std_df,
          file.path(wd, "output", "LM_standardized", "standardized_effect_sizes_all_panels.csv"),
          row.names = FALSE)

# Write one file per comparison x variable
for (pl in unique(std_df$panel)) {
  for (var in unique(std_df$variable)) {

    df_sub <- std_df %>%
      filter(panel == pl, variable == var)

    # skip empty combinations
    if (nrow(df_sub) == 0) next

    file_name <- paste0(
      "standardized_",
      gsub("[^A-Za-z0-9]+", "_", pl),
      "_",
      gsub("[^A-Za-z0-9]+", "_", var),
      ".csv"
    )

    write.csv(
      df_sub,
      file.path(wd, "output", "LM_standardized", file_name),
      row.names = FALSE
    )
  }
}
```

# 7. Sensitivity analysis for unobserved confounding (sensemakr)

## Function to run sensitivity analysis

```{r}
# Function to run sensitivity analysis + print Table X summary row
run_sensemakr <- function(
  model_object,
  comparison_name,
  wd
) {

  # Check available fitted models
  if (is.null(model_object$models)) {
    stop("model_object must contain a $models list")
  }

  uids <- names(model_object$models)

  # Create empty results data.frame
  sense_results <- data.frame(
    UID = character(),
    treatment_term = character(),
    r2yd_x = numeric(), # partial R² (treatment ~ outcome)
    rv_q = numeric(), # robustness value (q = 1)
    rv_qa = numeric(), # robustness value (q = 1, α = 0.05)
    benchmark_var = character(),
    r2yz_dx = numeric(), # strength of confounder with outcome
    r2dz_x = numeric(), # strength of confounder with treatment
    stringsAsFactors = FALSE
  )

  # Loop over each project (UID)
  for (uid in uids) {
    
    msg("──────────────────────────────────────────")
    msg("Processing REDD site: ", uid)

    model <- model_object$models[[uid]]

    if (is.null(model)) {
      sense_results <- rbind(sense_results, data.frame(
        UID = uid,
        treatment_term = NA,
        r2yd_x = NA,
        rv_q = NA,
        rv_qa = NA,
        benchmark_var = NA,
        r2yz_dx = NA,
        r2dz_x = NA
      ))
      next
    }

    # Identify treatment coefficient
    coef_mat <- summary(model)$coefficients
    coef_names <- rownames(coef_mat)

    # -----------------------------
    # Identify treatment term
    # -----------------------------
    # Treatment term to start with "group"
    group_terms <- coef_names[grepl("^group", coef_names)]

    # Require unique treatment term
    if (length(group_terms) != 1) {
      sense_results <- rbind(sense_results, data.frame(
        UID = uid,
        treatment_term = NA,
        r2yd_x = NA,
        rv_q = NA,
        rv_qa = NA,
        benchmark_var = NA,
        r2yz_dx = NA,
        r2dz_x = NA
      ))
      next
    }

    treatment_term <- group_terms[1]

    # -----------------------------
    # Identify benchmark covariate
    # -----------------------------
    coef_df <- as.data.frame(coef_mat)
    coef_df$term <- rownames(coef_df)

    # Keep only non-treatment, non-intercept, non-pair terms
    candidate_covars <- coef_df[
      !(coef_df$term %in% c("(Intercept)", treatment_term)) &
        !grepl("^factor\\(pair_id\\)", coef_df$term),
    ]

    # If no covariates available
    if (nrow(candidate_covars) == 0) {
      benchmark_var <- NA
      benchmark_t <- NA
    } else {
      # Select largest absolute t-value
      candidate_covars$abs_t <- abs(candidate_covars[, "t value"])

      best_idx <- which.max(candidate_covars$abs_t)

      benchmark_var <- candidate_covars$term[best_idx]
      benchmark_t <- candidate_covars[best_idx, "t value"]
    }

    # -----------------------------
    # Run sensemakr
    # -----------------------------
    sm <- tryCatch(
      sensemakr(
        model = model,
        treatment = treatment_term,
        benchmark_covariates = benchmark_var,
        kd = 1
      ),
      error = function(e) NULL
    )

    if (is.null(sm)) {
      sense_results <- rbind(sense_results, data.frame(
        UID = uid,
        treatment_term = treatment_term,
        r2yd_x = NA,
        rv_q = NA,
        rv_qa = NA,
        benchmark_var = benchmark_var,
        r2yz_dx = NA,
        r2dz_x = NA
      ))
      next
    }

    # Extract sensitivity stats (rounded %)
    r2yd_x <- round(sm$sensitivity_stats$r2yd.x * 100, 1)
    rv_q   <- round(sm$sensitivity_stats$rv_q * 100, 1)
    rv_qa  <- round(sm$sensitivity_stats$rv_qa * 100, 1)
    
    r2yz_dx <- round(sm$bounds$r2yz.dx * 100, 1)
    r2dz_x  <- round(sm$bounds$r2dz.x * 100, 1)
    
    sense_results <- rbind(sense_results, data.frame(
      UID = uid,
      treatment_term = treatment_term,
      r2yd_x = r2yd_x,
      rv_q = rv_q,
      rv_qa = rv_qa,
      benchmark_var = benchmark_var,
      r2yz_dx = r2yz_dx,
      r2dz_x = r2dz_x
    ))
  }

  # -----------------------------
  # Save results
  # -----------------------------
  model_name <- deparse(substitute(model_object))
  model_name <- sub("_[0-9]+$", "", model_name)

  out_path <- paste0(
    wd,
    "output/unobserved confounding/",
    comparison_name,
    "/",
    model_name,
    ".csv"
  )

  dir.create(dirname(out_path), recursive = TRUE, showWarnings = FALSE)
  write.csv(sense_results, out_path, row.names = FALSE)

  return(sense_results)
}
```

## Run sensitivity analysis

```{r}
# Comparisons:
# (1) 'REDD, No PA' vs 'No REDD, No PA'
# (2) 'REDD, No PA' vs 'No REDD, PA'
# (3) 'REDD, PA'    vs 'No REDD, No PA'
# (4) 'REDD, PA'    vs 'No REDD, PA'

# 7.1 REDD, No PA vs No REDD, No PA
sense_forestInt_1 <- run_sensemakr(
  model_object = forestInt_1,
  comparison_name = "REDD_noPA_vs_noREDD_noPA",
  wd = wd)

sum(!is.na(sense_carbon_flux_4$r2yd_x))
median(sense_carbon_flux_4$r2yd_x, na.rm = TRUE)
quantile(sense_carbon_flux_4$r2yd_x, 0.25, na.rm = TRUE)
quantile(sense_carbon_flux_4$r2yd_x, 0.75, na.rm = TRUE)
median(sense_carbon_flux_4$rv_q, na.rm = TRUE)
quantile(sense_carbon_flux_4$rv_q, 0.25, na.rm = TRUE)
quantile(sense_carbon_flux_4$rv_q, 0.75, na.rm = TRUE)
median(sense_carbon_flux_4$rv_qa, na.rm = TRUE)
quantile(sense_carbon_flux_4$rv_qa, 0.25, na.rm = TRUE)
quantile(sense_carbon_flux_4$rv_qa, 0.75, na.rm = TRUE)

sense_forestFrag_1 <- run_sensemakr(
  model_object = forestFrag_1,
  comparison_name = "REDD_noPA_vs_noREDD_noPA",
  wd = wd)

sense_biod_intact_1 <- run_sensemakr(
  model_object = biod_intact_1,
  comparison_name = "REDD_noPA_vs_noREDD_noPA",
  wd = wd)

sense_canopyheight_1 <- run_sensemakr(
  model_object = canopyheight_1,
  comparison_name = "REDD_noPA_vs_noREDD_noPA",
  wd = wd)

sense_carbon_flux_1 <- run_sensemakr(
  model_object = carbon_flux_1,
  comparison_name = "REDD_noPA_vs_noREDD_noPA",
  wd = wd)

# 7.2 REDD, No PA vs No REDD, PA
sense_forestInt_2 <- run_sensemakr(
  model_object = forestInt_2,
  comparison_name = "REDD_noPA_vs_noREDD_PA",
  wd = wd)

sense_forestFrag_2 <- run_sensemakr(
  model_object = forestFrag_2,
  comparison_name = "REDD_noPA_vs_noREDD_PA",
  wd = wd)

sense_biod_intact_2 <- run_sensemakr(
  model_object = biod_intact_2,
  comparison_name = "REDD_noPA_vs_noREDD_PA",
  wd = wd)

sense_canopyheight_2 <- run_sensemakr(
  model_object = canopyheight_2,
  comparison_name = "REDD_noPA_vs_noREDD_PA",
  wd = wd)

sense_carbon_flux_2 <- run_sensemakr(
  model_object = carbon_flux_2,
  comparison_name = "REDD_noPA_vs_noREDD_PA",
  wd = wd)

# 7.3 REDD, PA vs No REDD, No PA
sense_forestInt_3 <- run_sensemakr(
  model_object = forestInt_3,
  comparison_name = "REDD_PA_vs_noREDD_noPA",
  wd = wd)

sense_forestFrag_3 <- run_sensemakr(
  model_object = forestFrag_3,
  comparison_name = "REDD_PA_vs_noREDD_noPA",
  wd = wd)

sense_biod_intact_3 <- run_sensemakr(
  model_object = biod_intact_3,
  comparison_name = "REDD_PA_vs_noREDD_noPA",
  wd = wd)

sense_canopyheight_3 <- run_sensemakr(
  model_object = canopyheight_3,
  comparison_name = "REDD_PA_vs_noREDD_noPA",
  wd = wd)

sense_carbon_flux_3 <- run_sensemakr(
  model_object = carbon_flux_3,
  comparison_name = "REDD_PA_vs_noREDD_noPA",
  wd = wd)

# 7.4 REDD, PA vs No REDD, PA
sense_forestInt_4 <- run_sensemakr(
  model_object = forestInt_4,
  comparison_name = "REDD_PA_vs_noREDD_PA",
  wd = wd)

sense_forestFrag_4 <- run_sensemakr(
  model_object = forestFrag_4,
  comparison_name = "REDD_PA_vs_noREDD_PA",
  wd = wd)

sense_biod_intact_4 <- run_sensemakr(
  model_object = biod_intact_4,
  comparison_name = "REDD_PA_vs_noREDD_PA",
  wd = wd)

sense_canopyheight_4 <- run_sensemakr(
  model_object = canopyheight_4,
  comparison_name = "REDD_PA_vs_noREDD_PA",
  wd = wd)

sense_carbon_flux_4 <- run_sensemakr(
  model_object = carbon_flux_4,
  comparison_name = "REDD_PA_vs_noREDD_PA",
  wd = wd)
```
