---
title: "REDD_MatchingCode"
output: pdf_document
date: "2025-08-15"
---

Title: REDD+ Matching and Environmental Data Processing
Author: Kelly Ong

Description:
This script implements the spatial preprocessing and
propensity score matching pipeline for REDD+ project
polygons and their surrounding control areas.

Procedure:
1. Reads REDD+ and control polygons.
2. Loads spatial environmental datasets at multiple resolutions.
3. Masks and extracts data within an Area of Interest (AOI).
4. Selects matching covariates via coefficient of variation (CV) 
and variance inflation factor (VIF) filtering.
5. Performs nearest-neighbour matching with Mahalanobis distance.
6. Saves matched datasets, SMD statistics, and diagnostic plots.

# 0. Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)

# Set the working directory where all datasets are located
wd <- "~/Documents/Ecological Integrity of Avoided Deforestation projects/trial/data/"

# Load required packages
library(sf)         # vector data
library(terra)      # raster data
library(dplyr)      # data wrangling
library(ggplot2)    # plotting
library(MatchIt)    # matching
library(usdm)       # VIF (vifstep)
library(cobalt)     # balance diagnostics
library(tidyr)      # reshaping
library(ggspatial)  # map scalebars

# Output dirs
dirs_to_create <- c(
  "output/matching_REDD_and_control", 
  "output/NN_output_SMD_results_mahvars", 
  "output/NN_output_matched_data_mahvars",
  "output/NN_output_RDS_mahvars"
)
lapply(file.path(wd, dirs_to_create), dir.create, recursive = TRUE, showWarnings = FALSE)
```

# 1. Read core data

```{r}
# REDD+ project polygons (with REDD UID & biome info)
redd <- st_read(file.path(wd, "REDD.shp"))

# Control polygons (with REDD UID & biome info)
control <- st_read(file.path(wd, "REDD_control.shp"))

# Protected Areas raster (1 km resolution; cell value = 1 if PA present)
pa_raster <- rast(file.path(wd, "PA/pa_raster.tif"))

# Project-specific pixel resolution (km)
area <- read.csv(file.path(wd, "area.csv"))
```

# 2. Define resolutions and variable sets

```{r}
# Resolutions tested (in km)
resolutions <- c(1, 2, 4, 8, 16)

# Covariates for pixel-level propensity score matching
matching_datasets_names <- list(
  aboveground_biomass_2000 = "aboveground_biomass_2000",
  precipitation            = "precipitation",
  elevation                = "elevation",
  slope                    = "slope",
  human_footprint          = "human_footprint",
  gdp                      = "gdp",
  population_density       = "population_density",
  accessibility            = "accessibility"
)

# Ecological integrity datasets
env_datasets_names <- list(
  forestInt        = "forestInt",
  forestFrag       = "forestFrag",
  carbon_flux      = "carbon_flux",
  canopy_height    = "canopy_height",
  biodIntact2015   = "biodIntact2015"
)
```

# 3. Load rasters into nested lists

```{r}
# Each resolution is stored as a sublist of rasters for all variables.

load_rasters_by_res <- function(resolutions, var_names, base_folder) {
  out_list <- list()
  for (res in resolutions) {
    sublist <- list()
    for (var in var_names) {
      f <- file.path(wd, base_folder, paste0("resolution_", res, "km"),
                     paste0(var, "_res", res, "km.tif"))
      if (!file.exists(f)) {
        warning("File not found: ", f)
        next
      }
      sublist[[var]] <- rast(f)
    }
    out_list[[as.character(res)]] <- sublist
  }
  out_list
}

matching_vars <- names(matching_datasets_names)
env_vars <- names(env_datasets_names)

matching_rasters_list <- load_rasters_by_res(resolutions, matching_vars, "matching")
env_rasters_list <- load_rasters_by_res(resolutions, env_vars, "environmental")
```

# 4. Helper functions

```{r}
# Mask a raster to the extent of a vector polygon, reprojecting if necessary
mask_to_area <- function(raster_obj, buffer_vect) {
  if (crs(raster_obj) != crs(buffer_vect)) {
    raster_obj <- project(raster_obj, crs(buffer_vect))
  }
  mask(crop(raster_obj, buffer_vect), buffer_vect, touches = TRUE)
}

# Coefficient of Variation (CV)
compute_cv <- function(x) {
  m <- mean(x, na.rm = TRUE)
  if (m == 0) return(0)
  (sd(x, na.rm = TRUE) / m) * 100
}

# Select variables with CV above a given threshold
select_vars_by_cv <- function(data, var_names, cv_threshold = 15) {
  cv_vals <- sapply(data[var_names], compute_cv)
  intersect(names(cv_vals)[cv_vals > cv_threshold], names(data))
}
```

# 5. Matching loop

```{r}
# Loops over REDD+ projects, performs masking, covariate selection,
# propensity score matching, and saves results.

n <- nrow(redd)

for (i in seq_len(n)) {
  
  # Periodic memory cleanup
  if (i %% 10 == 0) { gc(); cat("...gc at i =", i, "\n") }
  
  # A. Identify REDD polygon and attributes
  redd_polygon <- redd[i, ]
  current_uid <- redd_polygon$UID
  redd_vect <- vect(redd_polygon)
  
  # B. Identify control polygon for the same biome (same UID & biome)
  control_poly <- control[control$UID == current_uid &
                          control$BIOME_NUM == redd_polygon$BIOME_NUM, ]
  control_vect <- vect(control_poly)
  
  # C. Combine REDD and control areas into AOI
  aoi_vect <- terra::union(redd_vect, control_vect) 
  
  # D. Save AOI plot for QC
  buffer_plot <- ggplot() +
    geom_sf(data = control_poly, fill = "lightblue", alpha = 0.5, color = "blue") +
    geom_sf(data = redd_polygon, fill = "green", color = "darkgreen") +
    ggtitle(paste0(current_uid, " REDD polygon & control")) +
    theme_minimal() +
    annotation_scale()
  ggsave(file.path(wd, "output/matching_REDD_and_control", paste0(current_uid, ".png")),
         plot = buffer_plot)
  
  # E. Resolution lookup
  fact <- as.numeric(area$Resolution[area$UID == current_uid])
  if (length(fact) == 0) next
  
  # F. Mask matching and environmental rasters to AOI
  matching_sublist <- matching_rasters_list[[as.character(fact)]]
  env_sublist <- env_rasters_list[[as.character(fact)]]
  if (is.null(matching_sublist) || is.null(env_sublist)) next
  
  masked_matching_rasters <- lapply(matching_sublist, mask_to_area, buffer_vect = aoi_vect)
  masked_env_rasters <- lapply(env_sublist, mask_to_area, buffer_vect = aoi_vect)
  
  match_df <- as.data.frame(rast(masked_matching_rasters), na.rm = TRUE, xy = TRUE)
  env_df <- as.data.frame(rast(masked_env_rasters), na.rm = FALSE, xy = TRUE)
  
  # Merge matching and environmental data by coordinates
  data_merged <- merge(env_df, match_df, by = c("x", "y"))
  if (nrow(data_merged) == 0) next
  
  # G. Mark REDD+ presence
  precip_redd <- mask_to_area(matching_sublist[["precipitation"]], redd_vect)
  inout <- terra::extract(precip_redd, data_merged[, c("x", "y")])
  data_merged$inside_REDD <- ifelse(is.na(inout[, 2]), 0, 1)
  
  # H. Mark protected area presence
  pa_extract <- terra::extract(pa_raster, data_merged[, c("x", "y")])
  data_merged$inside_PA <- ifelse(pa_extract[, 2] == 1, 1, 0)
  
  # I. Covariate selection: CV filter + VIF filter
  existing_vars <- intersect(names(matching_datasets_names), names(data_merged))
  selected_vars_cv <- select_vars_by_cv(data_merged, existing_vars, cv_threshold = 15)
  if (length(selected_vars_cv) == 0) next
  
  if (length(selected_vars_cv) > 1) {
    vif_result <- vifstep(data_merged[, selected_vars_cv, drop = FALSE], th = 10)
    selected_vars <- vif_result@results$Variables
  } else {
    selected_vars <- selected_vars_cv
  }
  
  # Remove incomplete cases
  data_matching <- data_merged[complete.cases(data_merged[, c("inside_REDD", selected_vars)]), ]
  if (nrow(data_matching) == 0) next
  
  # J. Propensity score matching (1:1 NN, Mahalanobis distance)
  m.out <- tryCatch({
    matchit(
      inside_REDD ~ ., 
      data = data_matching[, c("inside_REDD", selected_vars), drop = FALSE],
      method = "nearest",
      distance = "logit",
      mahvars = as.formula(paste("~", paste(selected_vars, collapse = "+"))),
      caliper = 0.25,
      ratio = 1,
      replace = FALSE
    )
  }, error = function(e) NULL)
  if (is.null(m.out)) next
  
  saveRDS(m.out, file.path(wd, "output/NN_output_RDS_mahvars",
                           paste0("NN_matched_data_", current_uid, "_biome_", redd_polygon$BIOME_NUM, ".rds")))
  
  # K. Extract matched dataset and append metadata
  matched_data <- match.data(m.out)
  if (nrow(matched_data) == 0) next
  
  matched_data <- as.data.frame(matched_data)
  matched_ids <- rownames(matched_data)
  
  # columns you want to bring back from original data
  extra_cols <- intersect(
    c("x", "y", names(env_datasets_names), names(selected_vars)),
    colnames(data_matching)
  )
  
  # bind them back using rownames
  matched_data <- cbind(
    matched_data,
    data_matching[matched_ids, extra_cols, drop = FALSE]
  )
  
  # remove duplicated columns if any
  matched_data <- matched_data[, !duplicated(colnames(matched_data)), drop = FALSE]
  
  # add metadata
  matched_data$country_id <- redd_polygon$country_id
  matched_data$REDD <- current_uid
  matched_data$BIOME_NUM <- redd_polygon$BIOME_NUM
  
  # extract PA values
  pts <- terra::vect(
    matched_data[, c("x", "y")],
    geom = c("x", "y"),
    crs = "EPSG:4326"
  )
  
  pa_extract_matched <- terra::extract(pa_raster, pts)
  matched_data$inside_protected_area <- ifelse(pa_extract_matched[, 2] == 1, 1, 0)
  
  # L. Assign pair IDs between "REDD" and "no REDD"
  if (!is.null(m.out) && !is.null(m.out$match.matrix)) {
    
    # 1) Extract the matching matrix
    match_matrix <- m.out$match.matrix
    
    # 2) Convert match_matrix -> long format
    pairs_df <- as.data.frame(match_matrix) %>%
      mutate(treated = rownames(match_matrix)) %>%
      pivot_longer(
        cols = -treated,             # pivot everything EXCEPT 'treated'
        names_to = "control_slot",
        values_to = "control"
      ) %>%
      filter(!is.na(control)) %>%
      mutate(
        treated = as.numeric(treated),
        control = as.numeric(control)
      )
    
    # 3) Create a simple ID for each matched pair
    pairs_df$pair_id <- seq_len(nrow(pairs_df))
    
    # 4) Reshape so each observation is in its own row (treated or control)
    pairs_long <- pairs_df %>%
      dplyr::select(pair_id, id = treated) %>%
      mutate(role = "treated") %>%
      bind_rows(
        pairs_df %>%
          dplyr::select(pair_id, id = control) %>%
          mutate(role = "control")
      )
    
    # 5) Merge pair information back into matched_data
    matched_data$original_id <- rownames(matched_data)
    pairs_long$id <- as.character(pairs_long$id)
    
    final_matched_data <- matched_data %>%
      left_join(pairs_long, by = c("original_id" = "id"))
    
    # 6) Write out final_matched_data (with pair info, x,y, inside_protected_area, etc.)
    columns_to_save <- c(
      "pair_id", "role", "x", "y", "country_id", "REDD", 
      "inside_REDD", "inside_protected_area", 
      names(env_datasets_names), names(matching_datasets_names)
    )
    columns_to_save <- intersect(columns_to_save, colnames(final_matched_data))
    
    output_file <- paste0(wd, "output/NN_output_matched_data_mahvars/NN_matched_data_",
                          current_uid, "_biome_", redd[i, ]$BIOME_NUM, ".csv")
    write.csv(final_matched_data[, columns_to_save, drop = FALSE],
              file = output_file, row.names = FALSE)
    
    cat("Wrote matched data with pairing info for REDD", current_uid, "resolution", fact, "km and biome", redd[i, ]$BIOME_NUM, "\n")
  }
  
  # M. Save standardised mean differences (SMDs) for covariates
  balance_summary <- bal.tab(m.out, un = TRUE)
  smd_df <- data.frame(
    Covariate = rownames(balance_summary$Balance),
    SMD = round(balance_summary$Balance$Diff.Adj, 4)
  )
  write.csv(smd_df, file.path(wd, "output/NN_output_SMD_results_mahvars",
                              paste0("NN_SMD_", current_uid, "_biome_", redd_polygon$BIOME_NUM, ".csv")),
            row.names = FALSE)
  
  cat("Processed:", current_uid, "\n")
}
```

